Assessing the Population-level Conservation Effects of Marine Protected Areas

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
add TMB plot back in

@@

```{r}

@@

```{r raw-synth, include = FALSE}

@@

@@

nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = 1999 + 1:length(abundance_hat)) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" ) %>% 
  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
  select(log_abundance_hat, year, classcode, targeted)

>>>>

```{r tmb-fit, fig.cap="Difference-in-difference results from mixed effect model fit to the individual transect observation level of the data."}
tmb_runs$did_plot[[1]] + 
scale_y_continuous(name = "Estimated MPA Effect")
```

@@

@@

```{r raw-synth, include = FALSE, eval = FALSE}

  consistent_classcodes <- used_data %>%
      group_by(classcode, region) %>%
      filter(any_seen) %>%
      summarise(n_years = n_distinct(year)) %>%
      ungroup() %>%
      filter(n_years == n_distinct(used_data$year))

synth_data <- used_data %>% 
  filter(classcode %in% consistent_classcodes$classcode,
         region %in% consistent_classcodes$region) %>% 
  mutate(classcode = ifelse(targeted == 1, "targeted", classcode),
         region = ifelse(targeted == 1, "targeted", region)) %>% 
      group_by(year,
               site_side,
               region,
               zone,
               transect,
               eventual_mpa,
               classcode,
               targeted) %>%
      summarise(
        total_classcode_density = sum(exp(log_density)),
        var_tex = sum(cumulative_n_obs),
        var_vis = mean(mean_vis),
        var_temp = mean(mean_temp),
        var_depth = mean(mean_depth),
        var_surge = mean(surge),
        var_kelp = mean(interp_kelp)
      ) %>%  # sum density across all levels of a transect
      group_by(year, site_side, region, eventual_mpa, classcode, targeted) %>%
      summarise(
        md = mean(total_classcode_density),
        var_tex = mean(var_tex),
        var_vis = mean(var_vis),
        var_temp = mean(var_temp),
        var_depth = mean(var_depth),
        var_surge = mean(var_surge),
        var_kelp = mean(var_kelp)
      ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
      ungroup()

 synth_data <- synth_data %>% 
     group_by(year,region,classcode,targeted) %>%
      summarise(
        total_biomass_density = (sum(md) / 1e6) * 10000,
        # calculate total and mean biomass densities across all species per year site side
        mean_biomass_density = log((mean(md) / 1e6) * 10000 + 1e-3),
        var_tex = mean(var_tex),
        var_vis = mean(var_vis),
        var_temp = mean(var_temp),
        var_kelp = mean(var_kelp)
      ) %>%
   ungroup() %>% 
   mutate(classcode = ifelse(targeted == 1, classcode, paste(classcode, region, sep = '_'))) %>% 
      mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
   as.data.frame()

prepped_synth_data <-
  Synth::dataprep(
    foo = synth_data,
    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"var")],
    predictors.op = "mean",
    dependent = "mean_biomass_density",
    unit.variable = "numeric_classcode",
    unit.names.variable = "classcode",
    time.variable = "year",
    treatment.identifier = "targeted",
    controls.identifier = unique(synth_data$classcode[synth_data$targeted == 0]),
    time.predictors.prior = years[years <= 2003],
    time.optimize.ssr = years[years <= 2003],
    time.plot = years
  )

@@

total_gap_plot <- gaps %>% 
  ggplot(aes(year, gap)) + 
  # geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_vline(aes(xintercept = 2003), color = "red")+
  geom_line() + 
    labs(x = "Year", y = "Standard deviations from synthetic control")

@@

nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = 1999 + 1:length(abundance_hat)) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" ) %>% 
  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
  select(log_abundance_hat, year, classcode, targeted)

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
add TMB plot back in

where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state. **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience. **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.  Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species). $\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA. $\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species. The shape and scale parameters of the Gamma distribution are estimated as well. See SI for further details of the estimation model. `r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands. As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.

@@

With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification. We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a]. We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results. Generalizations of Takens' theorem indicate that if two variables are part of the same dynamic system, their individual dynamics should reflect their relative causal influence. Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping"). CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species. Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (see SI for additional information in CCM testing). However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.

>>>>

where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state. **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience. **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.  Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species). $\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA. $\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species. The shape and scale parameters of the Gamma distribution are estimated as well. See SI for further details of the estimation model. `r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands. As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8. See SI for sensitivity analyses of various model structures and synthetic controls results.

@@

With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification. We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a]. We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results. Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Fig.SXX). However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
